---
title: "Biodiversity_Assignment"
author: "Jonathan Farr, Payton Arthur, Sammie Foley, and Jules Jones"
date: "10/31/2021"
output:
  pdf_document: default
  html_document: default
---


Setting up our packages
```{r setup, include=FALSE}
library(vegan)
library(tidyverse)
library(labdsv)
library(performance) 
library(LambertW)
quadrat = read.csv("quadrat_data.csv")
quadrat = quadrat[-c(21:997),]
biodiversity = read.csv("biodiversity.csv")
biodiversity = biodiversity[-c(357:979),]
perc = read.csv("percent_cover.csv")
```

Seaweed percentages
```{r}
seaweeds = perc %>%
    group_by(vertical_transect, quadrat) %>%
  mutate(unique_ID = as.numeric( paste0(vertical_transect, ".", quadrat))) %>%
  dplyr::select(c(unique_ID, quadrat, per_upright, per_crust)) %>%
  arrange(unique_ID)

seaweeds_clean = data.frame(seaweeds[,-c(1,3)])


# setting up our dataframe so that we can use matrify to arrange
#our data in long form and  then calculate diversity with vegan 
cleaned_biodiversity = biodiversity %>% # cleaning data to use matrify
    drop_na(n) %>%
    group_by(vertical_transect, quadrat) %>%
  mutate(unique_ID = as.numeric( paste0(vertical_transect, ".", quadrat))) %>%
  dplyr::select(unique_ID, species, n)

cleaned_biodiversity = data.frame(cleaned_biodiversity[,-c(1:2)]) # removing extra columns

# putting it into a matrix format
biodiversity_matrix <- matrify(cleaned_biodiversity) # using matrify

```

Calculating diversity indices and linking it to seaweed cover
```{r}

ShannonIndex <- data.frame(
  diversity(biodiversity_matrix, index = "shannon"))
head(ShannonIndex)
ShannonIndex = ShannonIndex %>%
  rename("ShannonIndex" = diversity.biodiversity_matrix..index....shannon..)

richness <- biodiversity %>% # calculating richness per quadrat
  drop_na(n) %>%
  group_by(vertical_transect,quadrat) %>%
  summarize(rich = length(unique(species)))   %>%
  mutate(unique_ID = as.numeric( paste0(vertical_transect, ".", quadrat)))

Data = cbind(ShannonIndex, seaweeds_clean, richness[,3]) # comining all our data
```

Now, we plot the relationship between diversity and seaweed cover using ggplot and geom_point
```{r}

# visualizing upright percent cover and diversity
p1 = ggplot(data = Data, aes(x=per_upright, y =ShannonIndex))+ 
  geom_point()+
  geom_smooth(method=lm, se=FALSE)

# visualizing crust percent cover and diversity
p2 = ggplot(data = Data, aes(x=per_crust, y =ShannonIndex))+ 
  geom_point()+
  geom_smooth(method=lm, se=FALSE)

# visualizing upright percent cover and richness
p3 = ggplot(data = Data, aes(x=per_upright, y =rich))+ 
  geom_point()+
  geom_smooth(method=lm, se=FALSE)

# visualizing crust percent cover and richness
p4 = ggplot(data = Data, aes(x=per_crust, y =rich))+ 
  geom_point()+
  geom_smooth(method=lm, se=FALSE)



```

```{r}
ggsave(plot = p1, filename = "upright_diversity.pdf")
ggsave(plot = p2, filename = "crustose_diversity.pdf")
ggsave(plot = p3, filename = "upright_richness.pdf")
ggsave(plot = p4, filename = "crustoes_richness.pdf")
```




Using regression to examine the relationships between seaweed % cover and shannon weiner diversity
We used poission distribution because it's more appropriate for data that's never negative
```{r}
# simple linear regression for upright and diversity
upright_diversity = lm(ShannonIndex ~ per_upright, data = Data)
summary(upright_diversity) # not significant

# simple linear regression for crust and diversity
crust_diversity = lm(ShannonIndex ~ per_crust, data = Data)
summary(crust_diversity) # not significant


# simple linear regression for upright and richness
upright_rich = lm(rich ~ per_upright, data = Data)
summary(upright_rich) # not significant

# simple linear regression for crust and richness
crust_rich = lm(rich ~ per_crust, data = Data)
summary(crust_rich) # not significant

# we have no significant results (teehee) but that's ok! science
```


```{r, echo=FALSE}


table = data.frame(Analysis = c("Shannon-Wiener Index", "Shannon-Wiener Index", "Richness", "Richness"), 
                   "Algae Type" = c("upright", "crustose", "upright", "crustose"), 
                   "P-Value" = c("0.1433", "0.6993", "0.2085", "0.9387"), 
                   "Adjusted R-squared" = c("0.06596", "-0.0466", "0.03562", "-0.0552"))

knitr::kable(table, caption= "Table 1. Linear regression results of the relationship between algae morphology and both Shannon-Wiener Index and Richness")



```

