---
title: "Biodiversity_Assignment"
author: "Jonathan Farr"
date: "10/31/2021"
output: html_document
---


Setting up our packages
```{r setup, include=FALSE}
library(vegan)
library(tidyverse)
library(labdsv)
library(performance) 
library(LambertW)
quadrat = read.csv("quadrat_data.csv")
quadrat = quadrat[-c(21:997),]
biodiversity = read.csv("biodiversity.csv")
biodiversity = biodiversity[-c(357:979),]
perc = read.csv("percent_cover.csv")
```

## Question

How does percent-cover of upright and crust seaweeds affect intertidal invertebrate diversity?

## Hypothesis

If an area has a higher percent cover of upright seaweeds, then intertidal diversity will be higher because they provide microhabitats and are recruiters.  

## Data Collection and Data Type

At Scott's Bay, we placed two transects parallel to the shoreline with one wave exposed and the other wave-sheltered. On each transect, we surveyed three locations using 3-4 quadrats per location to record the following: percent cover of upright algae, percent cover of crustose algae, and the species abundance of invertebrates (down to the lowest taxa possible). 

## Analysis Summary

We calculated the species richness and diversity (using the Shannon-Wiener index) of quadrates compared to percent cover of upright algae and crustose algae.  We then performed a simple linear regression to see if percent cover of crustose or upright algae has a significant effect on diversity and richness.

## Results

### Calculations
Seaweed percentages
```{r}
seaweeds = perc %>%
    group_by(vertical_transect, quadrat) %>%
  mutate(unique_ID = as.numeric( paste0(vertical_transect, ".", quadrat))) %>%
  dplyr::select(c(unique_ID, quadrat, per_upright, per_crust)) %>%
  arrange(unique_ID)

seaweeds_clean = data.frame(seaweeds[,-c(1,3)])


# setting up our dataframe so that we can use matrify to arrange
#our data in long form and  then calculate diversity with vegan 
cleaned_biodiversity = biodiversity %>% # cleaning data to use matrify
    drop_na(n) %>%
    group_by(vertical_transect, quadrat) %>%
  mutate(unique_ID = as.numeric( paste0(vertical_transect, ".", quadrat))) %>%
  dplyr::select(unique_ID, species, n)

cleaned_biodiversity = data.frame(cleaned_biodiversity[,-c(1:2)]) # removing extra columns

# putting it into a matrix format
biodiversity_matrix <- matrify(cleaned_biodiversity) # using matrify

```

Calculating diversity indices and linking it to seaweed cover
```{r}

ShannonIndex <- data.frame(
  diversity(biodiversity_matrix, index = "shannon"))
head(ShannonIndex)
ShannonIndex = ShannonIndex %>%
  rename("ShannonIndex" = diversity.biodiversity_matrix..index....shannon..)

richness <- biodiversity %>% # calculating richness per quadrat
  drop_na(n) %>%
  group_by(vertical_transect,quadrat) %>%
  summarize(rich = length(unique(species)))   %>%
  mutate(unique_ID = as.numeric( paste0(vertical_transect, ".", quadrat)))

Data = cbind(ShannonIndex, seaweeds_clean, richness[,3]) # comining all our data
```

Now, we plot the relationship between diversity and seaweed cover using ggplot and geom_point
```{r}

# visualizing upright percent cover and diversity
figure1<-ggplot(data = Data, aes(x=per_upright, y =ShannonIndex))+ 
  geom_point()+
  geom_smooth(method=lm, se=FALSE)+
  labs(x="Percent cover of upright algae (%)", y="Shannon-Wiener index")

# visualizing crust percent cover and diversity
figure2<-ggplot(data = Data, aes(x=per_crust, y =ShannonIndex))+ 
  geom_point()+
  geom_smooth(method=lm, se=FALSE)+
  labs(x="Percent cover of crustose algae (%)", y="Shannon-Wiener index")

# visualizing upright percent cover and richness
figure3<-ggplot(data = Data, aes(x=per_upright, y =rich))+ 
  geom_point()+
  geom_smooth(method=lm, se=FALSE)+
  labs(x="Percent cover of upright algae", y="Richness")

# visualizing crust percent cover and richness
figure4 <- ggplot(data = Data, aes(x=per_crust, y =rich))+ 
  geom_point()+
  geom_smooth(method=lm, se=FALSE)+
  labs(x="Percent cover of crustose algae", y="Richness")


figure1
```

Figure 1. The impact of percent cover of upright algae on the diversity of marine invertebrates (as calculated by the Shannon-Wiener Index)

```{r}
figure2
```

Figure 2. The impact of percent cover of crustose algae on the diversity of marine invertebrates (as calculated by the Shannon-Wiener Index)

```{r}
figure3
```

Figure 3. The impact of percent cover of crustose algae on the richness of marine invertebrates (as calculated by the Shannon-Wiener Index)

```{r}
figure4
```

Figure 4. The impact of percent cover of crustose algae on the richness of marine invertebrates (as calculated by the Shannon-Wiener Index)

### Linear Regression

Using regression to examine the relationships between seaweed % cover and shannon weiner diversity
We used poission distribution because it's more appropriate for data that's never negative
```{r}
# simple linear regression for upright and diversity
upright_diversity = lm(ShannonIndex ~ per_upright, data = Data)
summary(upright_diversity) # not significant

# simple linear regression for crust and diversity
crust_diversity = lm(ShannonIndex ~ per_crust, data = Data)
summary(crust_diversity) # not significant


# simple linear regression for upright and richness
upright_rich = lm(rich ~ per_upright, data = Data)
summary(upright_rich) # not significant

# simple linear regression for crust and richness
crust_rich = lm(rich ~ per_crust, data = Data)
summary(crust_rich) # not significant

# we have no significant results (teehee) but that's ok! science
```

## Conclusions

As none of the p-values are less than 0.05, there is no significant relationship between algae type and both diversity and richness.


